**Density Dependence**

### 2023-03-07 Rmd/report mac.Rmd
checked

**Deterministic projections at $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age**

@@

**Stochastic projections at $F_{MSY}$ for different scenarios**

@@

**Stochastic projections at all $F_{MSY}$'s from all scenarios**

@@

The current management of fish stocks is based on achieving the fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY). Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms. Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality. A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE). The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework. To do this, OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined. These OMs were then used to evaluate the consequences of including or not including density dependence in management advice.

@@

Estimating density-dependence from observations was only possible for mass-at-age and maturity age. Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).

In the next step, we applied a natural mortality-at-age vector derived from the Lorenzen approach which would mimick the VPA assessment with fixed M-at-age. The figure below compares the SAM assessment with the VPA and the VPAM assessment. A main difference is in the recruitment estimates. Using age-varying M tends to flatten the recruitment pattern. The vpaM model was used as the base case to be compared against the density-dependent simulations.

@@

* VPA Base Case, with varying M-at-age
* Density dependence in mass-at-age,
* Density dependence in mass, maturity-at-age,
* Density dependence in mass, maturity and M-at-age,

@@

The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers. The software is publicly available via Github (https://github.com/martinpastoors/ddmse). All data is derived from WGWIDE 2022 @ICES2022WGWIDE.

>>>>

**Deterministic projections at** $F_{MSY}$ for different scenarios: trends in weight, maturity and M-at-age

@@

**Stochastic projections at** $F_{MSY}$ for different scenarios

@@

**Stochastic projections at all** $F_{MSY}$'s from all scenarios

@@

The current management of fish stocks is based on achieving a fishing mortality ($F_{MSY}$) that is expected to achieve the maximum sustainable yield (MSY). Estimation of MSY generally only includes the stock recruitment relationship when considering density dependent population regulation mechanisms. Density dependence, however, may also operate in processes such as growth, sexual maturity, and natural mortality. A main aim of this study was to condition Operating Models (OMs) on alternative hypotheses about density dependence for use in Management Strategy Evaluation (MSE). The objective is to evaluate the robustness of the assessments conducted to support the ICES advice framework. To do this, OMs were developed that include density dependence in growth, maturity, natural mortality, individually and combined. These OMs were then used to evaluate the consequences of including or not including density dependence in management advice.

@@

Estimating density-dependence from observations was only possible for mass-at-age and maturity-at-qge. Density dependence in natural mortality can only be explored from theoretical considerations (because current assessment does not include age or time varying natural mortality).

In the next step, we applied a natural mortality-at-age vector based on [@lorenzen2000allometry] which would mimick the VPA assessment with fixed M-at-age. The figure below compares the SAM assessment with the VPA and the VPAM assessment. A main difference is in the recruitment estimates. Using age-varying M tends to flatten the recruitment pattern. The vpaM model was used as the base case to be compared against the density-dependent simulations.

@@

-   VPA Base Case, with varying M-at-age
-   Density dependence in mass-at-age,
-   Density dependence in mass, maturity-at-age,
-   Density dependence in mass, maturity and M-at-age,

@@

The quality of this analyses lives up to the quality standards that are required by ICES and so can potentially be used when providing advice to fisheries managers. The software is publicly available via Github (<https://github.com/martinpastoors/ddmse>). All data is derived from WGWIDE 2022 @ICES2022WGWIDE.
